---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# more_theme_easing_ideas

People have tried to make theming easier in ggplot2

- ggthemes: creating good bundled themes (also within ggplot2 theme_minimal, theme_bw etc)
- ggeasy: creating functions that match natural language
- https://evamaerey.github.io/ggplot2_grammar_guide/themes.html: creating comprehensive resourse for consultation
- 


<!-- badges: start -->
<!-- badges: end -->


```{r cars}
library(tidyverse)
gapminder::gapminder %>%  
  filter(year == 1992) %>%  
  ggplot() +  
  aes(x = gdpPercap, y = lifeExp) +  
  geom_point(alpha = .8, shape = 21,  
             fill = "white") +  
  aes(col = continent, fill = continent) +  
  geom_point(alpha = .3, shape = 21) +  
  aes(size = pop) +  
  scale_size(guide = F) +  
  labs(title = "Wealth and expected longevity\n in 1992") +  
  labs(subtitle = "Data from gapminder package in R") +  
  labs(x = "GDP per Capita") +  
  labs(y = "Life Expectency") +  # a tag is also an option
  labs(col = "Continent", fill = "Continent") +  
  labs(caption = "Vis: @EvaMaeRey with ggplot") ->  
g; g

g +  
  theme(plot.title.position = "plot") +  
  theme(plot.tag =  
          element_text(color = "plum4", size = 25)) +  
  theme(plot.title =  
          element_text(color = "plum4",  
                       size = 20,  
                       lineheight = 2.5)) +  
  theme(plot.subtitle = element_text(color = "plum4")) +  
  theme(axis.title.y =  
          element_text(color = "plum4",  
                       size = rel(3))) +  
  theme(axis.text.y.left = element_text(color = "plum4")) +  
  theme(legend.title =  
          element_text(color = "plum4", size = 22)) +  
  theme(legend.text = element_text(color = "plum4")) +  
  theme(axis.title.x = element_text(color = "plum4")) +  
  theme(axis.text.x = element_text(color = "plum4")) +  
  theme(plot.caption =  
          element_text(color = "plum4", size = 20)) +  
  theme(text = element_text(family = "Times")) +  
  theme(text = element_text(face = "bold.italic")) +  
  theme(text = element_text(  
    margin = margin(t = 3, r = 3, b = 3, l = 3,  
                    unit = "pt"))) +  
  theme(text = element_text(angle = 2)) +  
  theme(plot.title = element_text(hjust = .5)) +  
  theme(axis.title.x = element_text(hjust = 1))
```

# proposal

```{r, eval = F}
# instead of:
theme(axis.text.y.left = element_text(color = "plum4"))
# this?:
theme_axis_text_y(color = "plum4")
```

# implementation demonstration w/ text element


```{r, eval = T}
theme_axis_text_y_left <- function(family = NULL,
                              face = NULL,
                              colour = NULL,
                              size = NULL,
                              hjust = NULL,
                              vjust = NULL,
                              angle = NULL,
                              lineheight = NULL,
                              color = NULL,
                              margin = NULL,
                              debug = NULL,
                              inherit.blank = FALSE,
                              element_blank = FALSE){
  
         theme_element <- theme(axis.text.y.left = 
                  element_text(family = family,
                  face = face,
                  colour = colour,
                  size = size,
                  hjust = hjust,
                  vjust = vjust,
                  angle = angle,
                  lineheight = lineheight,
                  color = color,
                  margin = margin,
                  debug = debug,
                  inherit.blank = inherit.blank))
            
            if(element_blank){
              theme(axis.text.y = element_blank())
            }

return(theme_element)
  
}

theme_axis_text_y_left(color = "red")
```

```{r}
g + theme_axis_text_y_left(color = "red")
g + theme(axis.text.y.left = element_text(color = "red"))

```

# generalization

# categorize elements

```{r}
elements_type <- tribble(
~element, ~element_type,
"line", "line",
"rect", "rect",
"text", "text",
"title", "text",
"aspect.ratio", "aspect ratio of the panel",
"axis.title", "text", 
"axis.title.x", "text",
"axis.title.x.top", "text",
"axis.title.x.bottom", "text",
"axis.title.y", "text",
"axis.title.y.left", "text",
"axis.title.y.right", "text",
"axis.text", "text",
"axis.text.x", "text",
"axis.text.x.top", "text",
"axis.text.x.bottom", "text",
"axis.text.y", "text",
"axis.text.y.left", "text",
"axis.text.y.right", "text",
"axis.ticks", "line",
"axis.ticks.x", "line",
"axis.ticks.x.top", "line",
"axis.ticks.x.bottom", "line",
"axis.ticks.y", "line",
"axis.ticks.y.left", "line",
"axis.ticks.y.right", "line",
"axis.ticks.length", "line",
"axis.ticks.length.x", "line",
"axis.ticks.length.x.top", "line",
"axis.ticks.length.x.bottom", "line",
"axis.ticks.length.y", "line",
"axis.ticks.length.y.left", "line",
"axis.ticks.length.y.right", "line",
"axis.line", "line",
"axis.line.x", "line",
"axis.line.x.top", "line",
"axis.line.x.bottom", "line",
"axis.line.y", "line",
"axis.line.y.left", "line",
"axis.line.y.right", "line",
"legend.background", "rect",
"legend.margin", "margin",
"legend.spacing", "unit",
"legend.spacing.x", "unit",
"legend.spacing.y", "unit",
"legend.key", "rect",
"legend.key.size", "unit",
"legend.key.height", "unit",
"legend.key.width", "unit",
"legend.text", "text",
"legend.text.align", 'the position of legends ("none", "left", "right", "bottom", "top", or two-element numeric vector)',
"legend.title", "text",
"legend.title.align", "text",
"legend.position", 'the position of legends ("none", "left", "right", "bottom", "top", or two-element numeric vector)',
"legend.direction", 'layout of items in legends ("horizontal" or "vertical")',
"legend.justification", 'anchor point for positioning legend inside plot ("center" or two-element numeric vector) or the justification according to the plot area when positioned outside the plot',
"legend.box", "rect",
"legend.box.just", 'justification of each legend within the overall bounding box, when there are multiple legends ("top", "bottom", "left", or "right")',
"legend.box.margin", "margin",
"legend.box.background", "rect",
"legend.box.spacing", "unit",
"panel.background", "rect",
"panel.border", "rect",
"panel.spacing", "unit",
"panel.spacing.x", "unit",
"panel.spacing.y", "unit",
"panel.grid", "line",
"panel.grid.major", "line",
"panel.grid.minor", "line",
"panel.grid.major.x", "line",
"panel.grid.major.y", "line",
"panel.grid.minor.x", "line",
"panel.grid.minor.y", "line",
"panel.ontop", 'option to place the panel (background, gridlines) over the data layers (logical). Usually used with a transparent or blank panel.background',
"plot.background", "rect",
"plot.title", "text",
"plot.title.position", 'Alignment of the plot title/subtitle and caption. The setting for plot.title.position applies to both the title and the subtitle. A value of "panel" (the default) means that titles and/or caption are aligned to the plot panels. A value of "plot" means that titles and/or caption are aligned to the entire plot (minus any space for margins and plot tag).',
"plot.subtitle", "text",
"plot.caption", "text",
"plot.caption.position", 'Alignment of the plot title/subtitle and caption. The setting for plot.title.position applies to both the title and the subtitle. A value of "panel" (the default) means that titles and/or caption are aligned to the plot panels. A value of "plot" means that titles and/or caption are aligned to the entire plot (minus any space for margins and plot tag).',
"plot.tag", "text",
"plot.tag.position", 'The position of the tag as a string ("topleft", "top", "topright", "left", "right", "bottomleft", "bottom", "bottomright) or a coordinate. If a string, extra space will be added to accommodate the tag.',
"plot.margin", 'margin',
"strip.background", "rect",
"strip.background.x", "rect",
"strip.background.y", "rect",
"strip.clip", 'should strip background edges and strip labels be clipped to the extend of the strip background? Options are "on" to clip, "off" to disable clipping or "inherit" (default) to take the clipping setting from the parent viewport.',
"strip.placement", 'placement of strip with respect to axes, either "inside" or "outside". Only important when axes and strips are on the same side of the plot.',
"strip.text", "text",
"strip.text.x", "text",
"strip.text.x.bottom", "text",
"strip.text.x.top", "text",
"strip.text.y", "text",
"strip.text.y.left", "text",
"strip.text.y.right", "text",
"strip.switch.pad.grid", "unit",
"strip.switch.pad.wrap", "unit")

dim(elements_type)

elements_type %>% 
  count(element_type) %>% 
  filter(n >= 3)
  
```

# write batch function

## text elements

```{r}
text_theme_function_template <- 'theme_axis_text_y_left <- function(family = NULL,
                              face = NULL,
                              colour = NULL,
                              size = NULL,
                              hjust = NULL,
                              vjust = NULL,
                              angle = NULL,
                              lineheight = NULL,
                              color = NULL,
                              margin = NULL,
                              debug = NULL,
                              inherit.blank = FALSE,
                              element_blank = FALSE){
  
         theme_element <- theme(axis.text.y.left = 
                  element_text(family = family,
                  face = face,
                  colour = colour,
                  size = size,
                  hjust = hjust,
                  vjust = vjust,
                  angle = angle,
                  lineheight = lineheight,
                  color = color,
                  margin = margin,
                  debug = debug,
                  inherit.blank = inherit.blank)
                  )
            
            if(element_blank){
              theme_element <- theme(axis.text.y.left = 
              element_blank())
            }

return(theme_element)
  
}'


elements_type %>% 
  filter(element_type == "text") %>% 
  mutate(element_w_underscores = 
           element %>% str_replace_all("\\.", "_")) %>% 
  mutate(function_col = str_replace_all(text_theme_function_template, "axis_text_y_left", 
                                        element_w_underscores)) %>% 
  mutate(function_col = str_replace_all(function_col, "axis\\.text\\.y\\.left", 
                                        element)) ->
text_elements_type_functions
  
dim(text_elements_type_functions)

# inspect one of the prepared functions.
text_elements_type_functions[5,4][[1]] %>% cat

text_elements_type_functions$function_col %>% 
  paste(collapse = "
        
        
") %>% 
  writeLines(con = "R/text_elements_functions.R")
```

## line elements

```{r}
line_theme_function_template <- 'theme_axis_line_y_left <- function(
                              colour = NULL,
                              linewidth = NULL,
                                linetype = NULL,
                                lineend = NULL,
                                color = NULL,
                                arrow = NULL,
                                inherit.blank = FALSE,
                                size = NULL,
                                #size = deprecated(), # Error in deprecated() : could not find function "deprecated"
                                element_blank = FALSE){
  
         theme_element <- theme(axis.line.y.left = 
                  element_line(
                                colour = colour,
                                linewidth = linewidth,
                                linetype = linetype,
                                lineend = lineend,
                                color = color,
                                arrow = arrow,
                                inherit.blank = inherit.blank,
                                size = size
                              )
         )

            
            if(element_blank){
              theme_element <- theme(axis.line.y.left = 
              element_blank())
            }

return(theme_element)
  
}'


elements_type %>% 
  filter(element_type == "line") %>% 
  mutate(element_w_underscores = 
           element %>% str_replace_all("\\.", "_")) %>% 
  mutate(function_col = str_replace_all(line_theme_function_template, "axis_line_y_left", 
                                        element_w_underscores)) %>% 
  mutate(function_col = str_replace_all(function_col, "axis\\.line\\.y\\.left", 
                                        element)) ->
line_elements_type_functions
  
dim(line_elements_type_functions)

# inspect one of the prepared functions.
line_elements_type_functions[5,4][[1]] %>% cat

line_elements_type_functions$function_col %>% 
  paste(collapse = "
        
        
") %>% 
  writeLines(con = "R/line_elements_functions.R")
```


## rect elements

```{r}
rect_theme_function_template <- 'theme_panel_background <- function(
                                fill = NULL,
                                  colour = NULL,
                                  linewidth = NULL,
                                  linetype = NULL,
                                  color = NULL,
                                  inherit.blank = FALSE,
                                  size = NULL,
                                #size = deprecated(), # Error in deprecated() : could not find function "deprecated"
                                element_blank = FALSE){
  
         theme_element <- theme(panel.background = 
                  element_rect(
                                  fill = fill,
                                  colour = colour,
                                  linewidth = linewidth,
                                  linetype = linetype,
                                  color = color,
                                  inherit.blank = inherit.blank,
                                  size = size
                                )
         )

            
            if(element_blank){
              theme_element <- theme(panel.background = 
              element_blank())
            }

return(theme_element)
  
}'


elements_type %>% 
  filter(element_type == "rect") %>% 
  mutate(element_w_underscores = 
           element %>% str_replace_all("\\.", "_")) %>% 
  mutate(function_col = str_replace_all(rect_theme_function_template, "panel_background", 
                                        element_w_underscores)) %>% 
  mutate(function_col = str_replace_all(function_col, "panel\\.background", 
                                        element)) ->
rect_elements_type_functions
  
dim(rect_elements_type_functions)

# inspect one of the prepared functions.
rect_elements_type_functions[5,4][[1]] %>% cat

rect_elements_type_functions$function_col %>% 
  paste(collapse = "
        
        
") %>% 
  writeLines(con = "R/rect_elements_functions.R")
```


# trying some of this out

```{r}
source("R/text_elements_functions.R") # 30 proposed functions to adjust text.
source("R/line_elements_functions.R") # 29 proposed functions to adjust line elements
source("R/rect_elements_functions.R") # 7 proposed functions to adjust rect elements

g +  
  theme_axis_text_y_left(angle = 30, color = "red") +
  # theme_axis_text_x(angle = 50, hjust = 1) + 
  theme(axis.text.x = element_text(angle = 50, hjust = 1)) +
  # theme_text(color = "blue") + 
  theme(text = element_text(color = "blue")) + 
  theme_title(face = "bold.italic", lineheight = 1.5) + 
  theme_axis_line_x(colour = "darkred", 
                    linetype = "dashed", 
                    size = 2  # ggplot2_3.4.1  :-)
                    ) + 
  theme_panel_grid(color = "green") + 
  theme_panel_grid_minor(color = "orange") + 
  theme_panel_border(color = "magenta", size = 2, fill = alpha("pink", .2)) + 
  theme_legend_background(fill = "goldenrod1") + 
  theme_panel_background(fill = "yellow")

dim(elements_type)
```

# more to explore...

## margin & unit

```{r}
elements_type %>% 
  count(element_type) %>% 
  filter(n >= 3)

elements_type %>% 
  filter(element_type == "unit") 

elements_type %>% 
  filter(element_type == "margin") 

elements_type %>% 
  count(element_type) %>% 
  filter(n < 3) 


g +
  theme(legend.margin = margin(t = 1, unit='cm'))

g  + 
  theme(legend.box.spacing = unit(2, "cm"))
```
